Effect of temporal sampling

### 2019-10-24 R/temporal-sampling.Rmd
clean-up the Rmd from all the attempts

@@

We need to loop the different steps:<br>
  1)  Which samples are going to be sampled <br>
  2)  Run PCA on subset <br>
  3)  Run changepoint analysis on 1<sup>st</sup> (and maybe 2<sup>nd</sup>?) component <br>

```{r include=FALSE}
# Change here the sample size
sample_size <- c(rep(5:40, each=3),50)

output <- as.data.frame(matrix(rep(NA, 5*length(sample_size)), ncol=5))
colnames(output) <- c("Number_samples","PC1_var","PC2_var","AMOC_PC1","AMOC_PC2")

line=0
for (i in sample_size) {
  line=line+1
  # 1. Sample years
  which_samples <- sample(2:(nrow(zh)-1), size = sample_size[i],replace = F)
  which_samples <- c(which_samples,1,nrow(zh))
  which_samples <- which_samples[order(which_samples)]
  mdata <- zh[which_samples,]
  # reorder chronologically
  mdata <- mdata[order(mdata[,1]),]
  #remove rare species
  # present less than 3 times in the assemblage
  mdata <- mdata[,colSums(mdata>0)>3]

  # 2. Do ordination 
  PCAout<-decostand(mdata[,-1],"hellinger",na.rm=TRUE)
  PCAout<-dudi.pca(PCAout, scannf = FALSE)
  #summary(PCAout)

  # 3. Run changepoints on PC1
  cptmean.AMOC.PC1 <- cpt.mean(PCAout$li[,1], method="AMOC") 
  cptmean.AMOC.PC1 <- as.numeric(mdata[,1][cpts(cptmean.AMOC.PC1)])
  if(length(cptmean.AMOC.PC1)==0) cptmean.AMOC.PC1 <- NA

  # 4. Run changepoints on PC2
  cptmean.AMOC.PC2 <- cpt.mean(PCAout$li[,2], method="AMOC") 
  cptmean.AMOC.PC2 <- as.numeric(mdata[,1][cpts(cptmean.AMOC.PC2)])
  if(length(cptmean.AMOC.PC2)==0) cptmean.AMOC.PC2 <- NA

  # Save outputs
  output$Number_samples[line] <- i
  output$PC1_var[line]        <- PCAout$eig[1]/sum(PCAout$eig)*100
  output$PC2_var[line]        <- PCAout$eig[2]/sum(PCAout$eig)*100
  output$AMOC_PC1[line]       <- cptmean.AMOC.PC1
  output$AMOC_PC2[line]       <- cptmean.AMOC.PC2

}
output

# # PELT method with AIC penalty
# cptmean.AIC <- cpt.mean(y, method="PELT",penalty = "AIC")
# (cptmean.AIC.date<-as.numeric(time[cpts(cptmean.AIC)]))
# 
# # PELT method with manual penalty (elbow method)
# plot(0,0,xlim=c(0,20), ylim=c(0,15),col ="white", ylab="Number of changes in mean", xlab="Penalty", font=2,main="Elbow plot - choice of the penalty for changepoints analysis \nPruned Exact Linear Time (PELT) method")
# abline(h = c(seq(from=0, to=20, by=1),15), lty=3, col=adjustcolor("black",alpha.f = 0.5))
# abline(h=length(cptmean.AIC.date), lty=3, col="black", lwd=2)
# for (i in seq(from = 0, to = 20, by = 0.05)) {
#   cptmean=cpt.mean(y, method="PELT", penalty = "Manual", pen.value = i)
#   nb <- length(cpts(cptmean))
#   points(nb~i, pch=20)
# }
# legend("topright",legend = c("AIC","Manual - fluctuating penalty"), title = "Methods to chose the number of changes in mean",bty="n", pch = c(NA,20), lty=c(3,NA), lwd=c(2,NA), cex = 0.8)
# # change the pen.value according to the elbow plot
# cptmean.Manual <- cpt.mean(y, method="PELT",penalty = "Manual", pen.value = 20)
# (cptmean.Manual.date<-as.numeric(time[cpts(cptmean.Manual)]))
# 
# # Select here the final changepoint(s) you want to keep
# my_cpt <- (cptmean.Manual.date<-as.numeric(time[cpts(cptmean.Manual)]))
# if(is.null(my_cpt))
#   cat("\nWarning! No changepoints were detected. One of the condition for alternative stable state\nis an abrupt transition (see Andersen et al 2009, and our discussion in Bruel et al 2O18).\n\n")
# if(!is.null(my_cpt)) {
#   my_cpt <- as.data.frame(matrix(c(my_cpt,rep(NA,length(my_cpt))),ncol=2,byrow=F))
#   for (i in 1:nrow(my_cpt)) my_cpt[i,2]=time[which(time==my_cpt[i,1])+1]
#   # Re-arrange you my_cpt matrix
#   my_cpt <- my_cpt[order(my_cpt[,1],decreasing = T),order(my_cpt[1,],decreasing = T)]
# }

```

@@

```{r}
p1 <- ggplot(output, aes(AMOC_PC1, Number_samples)) + geom_point() + xlim(1850,2017)
p2 <- ggplot(output, aes(AMOC_PC2,Number_samples)) + geom_point() + xlim(1850,2017)
grid.arrange(p1,p2, nrow=2)
```
<br>`r fig_cap("timing of cpts", "Timing of changepoint")`

# Optimizing sampling by adding closest sample to shift

@@

```{r}
# Change here the sample size
sample_size <- c(rep(13:30, each=30))
#sample_size <- 15
output <- as.data.frame(matrix(rep(NA, 5*length(sample_size)), ncol=5))
colnames(output) <- c("Initial_number_samples","Initial_AMOC_PC1","Initial_AMOC_PC2","Number_needed_for_AMOC_PC1","Number_needed_for_AMOC_PC2")

output <- output[,c(1,2,4)]

for (i in 1:length(sample_size)) {
  if(i == 1)  {
    mdata <- zh[order(zh[,1]),]
    # 2. Do ordination 
    PCAout<-decostand(mdata[,-1],"hellinger",na.rm=TRUE)
    PCAout<-dudi.pca(PCAout, scannf = FALSE)
    true.eig1 <- PCAout$eig[1]/sum(PCAout$eig)*100
    true.eig2 <- PCAout$eig[2]/sum(PCAout$eig)*100

    # 3. Run changepoints on PC1
    true.cptmean.AMOC.PC1 <- cpt.mean(PCAout$li[,1], method="AMOC") 
    true.cptmean.AMOC.PC1 <- as.numeric(mdata[,1][cpts(true.cptmean.AMOC.PC1)])
    if(length(true.cptmean.AMOC.PC1)==0) true.cptmean.AMOC.PC1 <- NA

    # # 4. Run changepoints on PC2
    true.cptmean.AMOC.PC2 <- NA
    # true.cptmean.AMOC.PC2 <- cpt.mean(PCAout$li[,2], method="AMOC") 
    # true.cptmean.AMOC.PC2 <- as.numeric(mdata[,1][cpts(true.cptmean.AMOC.PC2)])
    # if(length(true.cptmean.AMOC.PC2)==0) true.cptmean.AMOC.PC2 <- NA

    # Save outputs
    c(nrow(zh), true.eig1, true.eig2,
      true.cptmean.AMOC.PC1, true.cptmean.AMOC.PC2)
  } 
  # 1. Sample years
  which_samples <- sample(2:(nrow(zh)-1), size = sample_size[i],replace = F)
  which_samples <- c(which_samples, 1, nrow(zh))
  which_samples <- which_samples[order(which_samples)]
  mdata <- zh[which_samples,]
  # reorder chronologically
  mdata <- mdata[order(mdata[,1]),]
  #remove rare species
  # present less than 3 times in the assemblage
  mdata <- mdata[,colSums(mdata>0)>3]

  # 2. Do ordination 
  PCAout<-decostand(mdata[,-1],"hellinger",na.rm=TRUE)
  PCAout<-dudi.pca(PCAout, scannf = FALSE)

  # 3. Run changepoints on PC1
  cptmean.AMOC.PC1 <- cpt.mean(PCAout$li[,1], method="AMOC") 
  cptmean.AMOC.PC1_num <- cptmean.AMOC.PC1
  cptmean.AMOC.PC1 <- as.numeric(mdata[,1][cpts(cptmean.AMOC.PC1)])
  if(length(cptmean.AMOC.PC1)==0) cptmean.AMOC.PC1 <- NA

  # # 4. Run changepoints on PC2
  # cptmean.AMOC.PC2 <- cpt.mean(PCAout$li[,2], method="AMOC") 
  # cptmean.AMOC.PC2_num <- cptmean.AMOC.PC2
  # cptmean.AMOC.PC2 <- as.numeric(mdata[,1][cpts(cptmean.AMOC.PC2)])
  # if(length(cptmean.AMOC.PC2)==0) cptmean.AMOC.PC2 <- NA

  # Save outputs
  output$Initial_number_samples[i]     <- sample_size[i]
  output$Initial_AMOC_PC1[i]           <- cptmean.AMOC.PC1
  #output$Initial_AMOC_PC2[i]           <- cptmean.AMOC.PC2
  if(cptmean.AMOC.PC1==true.cptmean.AMOC.PC1){
        output$Number_needed_for_AMOC_PC1[i] <- length(which_samples)
        numberPC1_ok = TRUE
        } else numberPC1_ok = FALSE
  # if(!is.na(cptmean.AMOC.PC2) & !is.na(true.cptmean.AMOC.PC2) & cptmean.AMOC.PC2==true.cptmean.AMOC.PC2) {
  #   output$Number_needed_for_AMOC_PC2[i] <- length(which_samples) 
  #   numberPC2_ok = TRUE
  # } else numberPC2_ok = FALSE

  # 5. Now, how many more samples are needed to get true cptmean.AMOC.PC1?
  for (j in 1:(nrow(zh)-sample_size[i])) {
    if(!numberPC1_ok) {
      h=1
      while((which_samples[cpts(cptmean.AMOC.PC1_num)]-h) %in% which_samples & which_samples[cpts(cptmean.AMOC.PC1_num)]-h > 0) h=h+1
      which_samples <- c(which_samples, which_samples[cpts(cptmean.AMOC.PC1_num)]-h)

      h=1
      while((which_samples[cpts(cptmean.AMOC.PC1_num)]+h) %in% which_samples & which_samples[cpts(cptmean.AMOC.PC1_num)]+h<nrow(zh)) h=h+1
      which_samples <- c(which_samples, which_samples[cpts(cptmean.AMOC.PC1_num)]+h)

      which_samples <- which_samples[which_samples>0 & which_samples<= nrow(zh)]
      which_samples <- unique(which_samples)

    which_samples <- which_samples[order(which_samples)]
    mdata <- zh[which_samples,]
    # reorder chronologically
    mdata <- mdata[order(mdata[,1]),]

    # 2. Do ordination 
    PCAout<-decostand(mdata[,-1],"hellinger",na.rm=TRUE)
    PCAout<-dudi.pca(PCAout, scannf = FALSE)
    #summary(PCAout)

    # 3. Run changepoints on PC1
    cptmean.AMOC.PC1 <- cpt.mean(PCAout$li[,1], method="AMOC") 
    cptmean.AMOC.PC1_num <- cptmean.AMOC.PC1
    cptmean.AMOC.PC1 <- as.numeric(mdata[,1][cpts(cptmean.AMOC.PC1)])
    if(length(cptmean.AMOC.PC1)==0) cptmean.AMOC.PC1 <- NA

    # # 4. Run changepoints on PC2
    # cptmean.AMOC.PC2 <- cpt.mean(PCAout$li[,2], method="AMOC") 
    # cptmean.AMOC.PC2_num <- cptmean.AMOC.PC2
    # cptmean.AMOC.PC2 <- as.numeric(mdata[,1][cpts(cptmean.AMOC.PC2)])
    # if(length(cptmean.AMOC.PC2)==0) cptmean.AMOC.PC2 <- NA

    if(!is.na(cptmean.AMOC.PC1) & !is.na(true.cptmean.AMOC.PC1) & cptmean.AMOC.PC1==true.cptmean.AMOC.PC1){
        output$Number_needed_for_AMOC_PC1[i] <- length(which_samples)
        numberPC1_ok = TRUE
        } else numberPC1_ok = FALSE
    # if(!is.na(cptmean.AMOC.PC2) & !is.na(true.cptmean.AMOC.PC2) & cptmean.AMOC.PC2==true.cptmean.AMOC.PC2) {
    #   output$Number_needed_for_AMOC_PC2[i] <- length(which_samples) 
    #   numberPC2_ok = TRUE
    # } else numberPC2_ok = FALSE
  }  
    }

}
output$total_added <- output$Number_needed_for_AMOC_PC1-output$Initial_number_samples-2
output
```

```{r}
ggplot(output, aes(Initial_number_samples,total_added)) + geom_point(alpha=.3)

```
<br>`r fig_cap("how many samples must be added to get true changepoint", "A 'biased' sampling (trying to detect the true shift by getting closest sample) is not a guarantee to reduce the total number of samples that should be analyzed to detect the true shift")`

# Optimizing sampling by adding intermediate samples between shift and next sample analyzed

```{r}
# Change here the sample size
sample_size <- c(rep(15:30, each=30))
#sample_size <- 15
output <- as.data.frame(matrix(rep(NA, 5*length(sample_size)), ncol=5))
colnames(output) <- c("Initial_number_samples","Initial_AMOC_PC1","Initial_AMOC_PC2","Number_needed_for_AMOC_PC1","Number_needed_for_AMOC_PC2")

output <- output[,c(1,2,4)]

for (i in 1:length(sample_size)) {
  if(i == 1)  {
    mdata <- zh[order(zh[,1]),]
    # 2. Do ordination 
    PCAout<-decostand(mdata[,-1],"hellinger",na.rm=TRUE)
    PCAout<-dudi.pca(PCAout, scannf = FALSE)
    true.eig1 <- PCAout$eig[1]/sum(PCAout$eig)*100
    true.eig2 <- PCAout$eig[2]/sum(PCAout$eig)*100

    # 3. Run changepoints on PC1
    true.cptmean.AMOC.PC1 <- cpt.mean(PCAout$li[,1], method="AMOC") 
    true.cptmean.AMOC.PC1 <- as.numeric(mdata[,1][cpts(true.cptmean.AMOC.PC1)])
    if(length(true.cptmean.AMOC.PC1)==0) true.cptmean.AMOC.PC1 <- NA

    # # 4. Run changepoints on PC2
    # true.cptmean.AMOC.PC2 <- cpt.mean(PCAout$li[,2], method="AMOC") 
    # true.cptmean.AMOC.PC2 <- as.numeric(mdata[,1][cpts(true.cptmean.AMOC.PC2)])
    # if(length(true.cptmean.AMOC.PC2)==0) true.cptmean.AMOC.PC2 <- NA

    # Save outputs
    c(nrow(zh), true.eig1, true.eig2,
      true.cptmean.AMOC.PC1, true.cptmean.AMOC.PC2)
  } 
  # 1. Sample years
  which_samples <- sample(2:(nrow(zh)-1), size = sample_size[i],replace = F)
  which_samples <- c(which_samples,1,nrow(zh))
  which_samples <- which_samples[order(which_samples)]
  mdata <- zh[which_samples,]
  # reorder chronologically
  mdata <- mdata[order(mdata[,1]),]
  #remove rare species
  # present less than 3 times in the assemblage
  mdata <- mdata[,colSums(mdata>0)>3]

  # 2. Do ordination 
  PCAout<-decostand(mdata[,-1],"hellinger",na.rm=TRUE)
  PCAout<-dudi.pca(PCAout, scannf = FALSE)

  # 3. Run changepoints on PC1
  cptmean.AMOC.PC1 <- cpt.mean(PCAout$li[,1], method="AMOC") 
  cptmean.AMOC.PC1_num <- cptmean.AMOC.PC1
  cptmean.AMOC.PC1 <- as.numeric(mdata[,1][cpts(cptmean.AMOC.PC1)])
  if(length(cptmean.AMOC.PC1)==0) cptmean.AMOC.PC1 <- NA

  # # 4. Run changepoints on PC2
  # cptmean.AMOC.PC2 <- cpt.mean(PCAout$li[,2], method="AMOC") 
  # cptmean.AMOC.PC2_num <- cptmean.AMOC.PC2
  # cptmean.AMOC.PC2 <- as.numeric(mdata[,1][cpts(cptmean.AMOC.PC2)])
  # if(length(cptmean.AMOC.PC2)==0) cptmean.AMOC.PC2 <- NA

  # Save outputs
  output$Initial_number_samples[i]     <- sample_size[i]
  output$Initial_AMOC_PC1[i]           <- cptmean.AMOC.PC1
  #output$Initial_AMOC_PC2[i]           <- cptmean.AMOC.PC2
  if(cptmean.AMOC.PC1==true.cptmean.AMOC.PC1){
        output$Number_needed_for_AMOC_PC1[i] <- length(which_samples)
        numberPC1_ok = TRUE
        } else numberPC1_ok = FALSE
  # if(!is.na(cptmean.AMOC.PC2) & !is.na(true.cptmean.AMOC.PC2) & cptmean.AMOC.PC2==true.cptmean.AMOC.PC2) {
  #   output$Number_needed_for_AMOC_PC2[i] <- length(which_samples) 
  #   numberPC2_ok = TRUE
  # } else numberPC2_ok = FALSE

  # 5. Now, how many more samples are needed to get true cptmean.AMOC.PC1?
  for (j in 1:(nrow(zh)-sample_size[i])) {
    if(!numberPC1_ok) {

      h=0
      max_samp <- round((which_samples[cpts(cptmean.AMOC.PC1_num)+h]+which_samples[cpts(cptmean.AMOC.PC1_num)+h+1])/2)
      while(max_samp %in% which_samples & (cpts(cptmean.AMOC.PC1_num)+h+1)<=nrow(zh)) {
        h=h+1
        other_limit <- ifelse(length(which_samples[cpts(cptmean.AMOC.PC1_num)+h+1])<=nrow(zh),
                              which_samples[cpts(cptmean.AMOC.PC1_num)+h+1],
                              nrow(zh))
        max_samp <- round((which_samples[cpts(cptmean.AMOC.PC1_num)+h]+which_samples[cpts(cptmean.AMOC.PC1_num)+h+1])/2)
        }

      h=0
      min_samp <- round((which_samples[cpts(cptmean.AMOC.PC1_num)-h]+which_samples[cpts(cptmean.AMOC.PC1_num)-h-1])/2)
      while(min_samp %in% which_samples & (cpts(cptmean.AMOC.PC1_num)-h-1)>0) {
        h=h+1
        other_limit <- ifelse(length(which_samples[cpts(cptmean.AMOC.PC1_num)-h-1])>0,
                              which_samples[cpts(cptmean.AMOC.PC1_num)-h-1],
                              1)
        min_samp <- round((which_samples[cpts(cptmean.AMOC.PC1_num)-h]+other_limit)/2)
      }

      which_samples <- c(which_samples, max_samp, min_samp)
      which_samples <- which_samples[!is.na(which_samples)]
      which_samples <- which_samples[which_samples>0 & which_samples<= nrow(zh)]
      which_samples <- unique(which_samples)

    which_samples <- which_samples[order(which_samples)]
    mdata <- zh[which_samples,]
    # reorder chronologically
    mdata <- mdata[order(mdata[,1]),]

    # 2. Do ordination 
    PCAout<-decostand(mdata[,-1],"hellinger",na.rm=TRUE)
    PCAout<-dudi.pca(PCAout, scannf = FALSE)
    #summary(PCAout)

    # 3. Run changepoints on PC1
    cptmean.AMOC.PC1 <- cpt.mean(PCAout$li[,1], method="AMOC") 
    cptmean.AMOC.PC1_num <- cptmean.AMOC.PC1
    cptmean.AMOC.PC1 <- as.numeric(mdata[,1][cpts(cptmean.AMOC.PC1)])
    if(length(cptmean.AMOC.PC1)==0) cptmean.AMOC.PC1 <- NA

    # # 4. Run changepoints on PC2
    # cptmean.AMOC.PC2 <- cpt.mean(PCAout$li[,2], method="AMOC") 
    # cptmean.AMOC.PC2_num <- cptmean.AMOC.PC2
    # cptmean.AMOC.PC2 <- as.numeric(mdata[,1][cpts(cptmean.AMOC.PC2)])
    # if(length(cptmean.AMOC.PC2)==0) cptmean.AMOC.PC2 <- NA

    if(!is.na(cptmean.AMOC.PC1) & !is.na(true.cptmean.AMOC.PC1) & cptmean.AMOC.PC1==true.cptmean.AMOC.PC1){
        output$Number_needed_for_AMOC_PC1[i] <- length(which_samples)
        numberPC1_ok = TRUE
        } else numberPC1_ok = FALSE
    # if(!is.na(cptmean.AMOC.PC2) & !is.na(true.cptmean.AMOC.PC2) & cptmean.AMOC.PC2==true.cptmean.AMOC.PC2) {
    #   output$Number_needed_for_AMOC_PC2[i] <- length(which_samples) 
    #   numberPC2_ok = TRUE
    # } else numberPC2_ok = FALSE
  }  
    }

}
output$total_added <- output$Number_needed_for_AMOC_PC1-output$Initial_number_samples
output
```

```{r}
ggplot(output, aes(Initial_number_samples,total_added)) + geom_point()
```
<br>`r fig_cap("how many samples must be added to get true changepoint method 2", "Adding progressively samples to fill in the gaps between transition and next/previous sample is a better strategy.")`

```{r}
output2 <- output %>% 
  group_by(Initial_number_samples) %>% 
  summarise(total0 = sum(total_added==0))

plot(output2)
abline(h=80)
```

# Chances of getting close to the real break

Following discussion with Easton on 10/10/2019.

```{r}
# Change here the threshold year: how close you need to get close to the actual year
threshold_yr = c(2,4,6,8,10,12,14)

# Change here the sample size
sample_size <- c(rep(15:50, each=100))
#sample_size <- 15
output3 <- as.data.frame(matrix(rep(NA, 4*length(sample_size)*length(threshold_yr)), ncol=4))
colnames(output3) <- c("Initial_number_samples","threshold_yr","Initial_AMOC_PC1","Within_threshold")

for (i in 1:length(sample_size)) {
  if(i == 1)  {
    row=0
    mdata <- zh[order(zh[,1]),]
    # 2. Do ordination 
    PCAout<-decostand(mdata[,-1],"hellinger",na.rm=TRUE)
    PCAout<-dudi.pca(PCAout, scannf = FALSE)
    true.eig1 <- PCAout$eig[1]/sum(PCAout$eig)*100
    true.eig2 <- PCAout$eig[2]/sum(PCAout$eig)*100

    # 3. Run changepoints on PC1
    true.cptmean.AMOC.PC1 <- cpt.mean(PCAout$li[,1], method="AMOC") 
    true.cptmean.AMOC.PC1 <- as.numeric(mdata[,1][cpts(true.cptmean.AMOC.PC1)])
    if(length(true.cptmean.AMOC.PC1)==0) true.cptmean.AMOC.PC1 <- NA

    # # 4. Run changepoints on PC2
    # true.cptmean.AMOC.PC2 <- cpt.mean(PCAout$li[,2], method="AMOC") 
    # true.cptmean.AMOC.PC2 <- as.numeric(mdata[,1][cpts(true.cptmean.AMOC.PC2)])
    # if(length(true.cptmean.AMOC.PC2)==0) true.cptmean.AMOC.PC2 <- NA

    # Save outputs
    c(nrow(zh), true.eig1, true.eig2,
      true.cptmean.AMOC.PC1, true.cptmean.AMOC.PC2)
  }

  for (j in 1:length(threshold_yr)) {
    row=row+1
      # 1. Sample years
  which_samples <- sample(2:(nrow(zh)-1), size = sample_size[i],replace = F)
  which_samples <- c(which_samples,1,nrow(zh))
  which_samples <- which_samples[order(which_samples)]
  mdata <- zh[which_samples,]
  # reorder chronologically
  mdata <- mdata[order(mdata[,1]),]
  #remove rare species
  # present less than 3 times in the assemblage
  mdata <- mdata[,colSums(mdata>0)>3]

  # 2. Do ordination 
  PCAout<-decostand(mdata[,-1],"hellinger",na.rm=TRUE)
  PCAout<-dudi.pca(PCAout, scannf = FALSE)

  # 3. Run changepoints on PC1
  cptmean.AMOC.PC1 <- cpt.mean(PCAout$li[,1], method="AMOC") 
  cptmean.AMOC.PC1_num <- cptmean.AMOC.PC1
  cptmean.AMOC.PC1 <- as.numeric(mdata[,1][cpts(cptmean.AMOC.PC1)])
  if(length(cptmean.AMOC.PC1)==0) cptmean.AMOC.PC1 <- 0

  # Save outputs
  output3$Initial_number_samples[row]     <- sample_size[i]
  output3$threshold_yr[row]               <- threshold_yr[j]
  output3$Initial_AMOC_PC1[row]           <- cptmean.AMOC.PC1
  if(abs(cptmean.AMOC.PC1-true.cptmean.AMOC.PC1)<=threshold_yr[j]){
        numberPC1_ok = 1
        } else {numberPC1_ok = 0}

  output3$Within_threshold[row] <- numberPC1_ok
  }

}
head(output3)
tail(output3)
```

```{r summary }
names(output3)

outsumm <- as.data.frame(with(output3, tapply(Within_threshold,list("Initial_number_samples"=Initial_number_samples,"threshold_yr"=threshold_yr), sum)))
outsumm$InitialNumber <- as.numeric(rownames(outsumm))
outsumm <- melt(outsumm, id.vars = "InitialNumber")
outsumm$more80 <- ifelse(outsumm$value>=80,1,0)
ggplot(outsumm, aes(InitialNumber, variable, fill=more80)) + geom_tile() +
  xlab("Sample analyzed") + ylab("Tolerance for error (year)")

@@

>>>>

**Note for next time: write function to compute the tolerance to error against sample size**

@@

@@

# Types of sampling

1. Random <br>
2. Regular <br>
3. Convergent <br>

```{r}
source("ar1_process.R")
source("R/uncertainty_tolerance.R")
args(ut)
```

@@

```{r, cache = T}
output <- ut(y=exp(ar1), sample_size = c(20:50), threshold_error = c(2:10))
names(output)
head(output$convergence_cptmean)
ggplot(output$convergence_cptmean, aes(sample_size,number_sample_added)) + geom_point(alpha=.3)

```{r summary }
names(output$random_cptmean)

outsumm <- as.data.frame(with(output$random_cptmean, tapply(same_cptmean,list("Initial_number_samples"=sample_size,"threshold_yr"=threshold_error), sum)))
outsumm$InitialNumber <- as.numeric(rownames(outsumm))
outsumm <- melt(outsumm, id.vars = "InitialNumber")
outsumm$more80 <- ifelse(outsumm$value>=7,1,0)
ggplot(outsumm, aes(InitialNumber, variable, fill=more80)) + geom_tile() +
  xlab("Sample analyzed") + ylab("Tolerance for error (year)")

@@

If we know a priori the shape of the relationship, a non-random sampling is preferred, e.g., <a href="https://onlinelibrary.wiley.com/doi/full/10.1002/env.2552"> VB growth function </a>.
